### 2022-10-01 Extra.tex
Progress

 U.S. attention to internet freedom continued to grow throughout the 2010's, with Secretary of State Hillary Clinton calling for internet freedom abroad (\cite{government2010}) and President Obama defending a version of it at home (\cite{government2016}) and overseas (\cite{richburg2009}).

 DiD? Similar levels of internet freedom in the past, then Chinese aid begins to increase, then what happens to U.S. aid? Or backwards? (U.S. aid trends and relatively similar (parallel trends) then Chinese aid begins) (OR look at two countries with different internet freedom levels that have received relatively similar levels of US Aid, then see what happens after Chinese aid begins) if no DiD, I can definitely do this as a case study.